# Reproducible Research: Peer Assessment 1
For this assignment the data  was processed using R version 3.1.0 in linux Ubuntu 12.04

## Loading and preprocessing the data
The first pass which was taken to load the data in the work directory was to extract them from the zip file and then was read it using the read.csv command to stored them in a object called "activity" and the first preprocessing step was to convert  the variable "date" in to a date type variable.
```{r}
unzip("activity.zip")
activity<-read.csv("activity.csv")
#processing the data
#turn the date variable into a date type variable 
activity$date<-as.Date(activity$date)
```

## What is mean total number of steps taken per day?
To answer this question was created a vector variable called "totalsteps" in which  every element is the sum  of  the steps taken each day, to do that  was used the tapply function to apply the function sum on the steps by date as factor variable.
```{r}
#total of steps taken per day 
totalsteps<-with(activity,tapply(steps,as.factor(date),sum))
# Histogram of the total number of steps taken each day
hist(totalsteps, main="Total Number of Steps Taken Each Day",xlab="Total steps",col="blue")
# the mean and median total number of steps taken per day
#Mean of the total steps
tmean<-mean(totalsteps,na.rm=T)
#Median of the total steps
tmedian<-median(totalsteps,na.rm=T)
#setting inline output options to avoid the default scientific notation
options(scipen = 1, digits = 2)
```

The mean of the total steps taken per day is `r tmean` and its median is `r tmedian`.

## What is the average daily activity pattern?
```{r}
#time series plot
averageInterval<-aggregate(steps~interval,data=activity,mean)
with(averageInterval,plot(interval,steps,type="l"))
title("Average steps taken across the 5-minutes interval")
#maximum number steps
averageInterval[averageInterval$steps==max(averageInterval$steps),]
```



## Imputing missing values
```{r}
#4.1 total number of missing values in the dataset 
sum(is.na(activity))

#4.2 filling missindg data

#4.3 data set without missing values
activity_no_missing<-activity
for(i in 1:length(activity_no_missing[,1])){
    if(is.na(activity_no_missing[i,1])){
        b<-activity_no_missing[activity_no_missing[,3]==activity_no_missing[i,3],]
        activity_no_missing[i,1]<-round(mean(b[,1],na.rm=T))
        }
    }

#4.4 Histogram
#total of steps taken per day 
totalsteps2<-with(activity_no_missing,tapply(steps,as.factor(date),sum))
# 2.1 Histogram of the total number of steps taken each day
hist(totalsteps2, main="Total Number of Steps Taken Each Day",xlab="Total steps",col="red")
# 2.2 the mean and median total number of steps taken per day
#Mean
mean(totalsteps2,na.rm=T)
#Median
median(totalsteps2,na.rm=T)
```



## Are there differences in activity patterns between weekdays and weekends?
```{r}
for(i in 1:length(activity_no_missing[,1])){
    if(!(weekdays(activity_no_missing[i,2]) %in% c("sábado","domingo"))){
        activity_no_missing[i,4]<-"weekday"
        }else{
            activity_no_missing[i,4]<-"weekend"
            }
    }

activity_no_missing[,4]<-as.factor(activity_no_missing[,4])
colnames(activity_no_missing)[4]<-"weekdayOrWeekend"

averageIntervalWeek<-aggregate(steps~interval+weekdayOrWeekend,data=activity_no_missing,mean)

library(lattice)
with(averageIntervalWeek,xyplot(round(steps)~interval|weekdayOrWeekend,layout=c(1,2),type="l"))

```

